A comparative analysis of penalized linear mixed models in structured genetic data

### 2020-09-24 main.tex
3.5 small tweaks

Furthermore, the proportion of the environmental effect in the column space of $\boldsymbol{P}_{\bX}$ (i.e., the portion that will cause confounding) is given by
\begin{equation}
    \label{eqn:ratio}
    ||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2,
\end{equation}
which we refer to as the Confounding Proportion (CP). The CP can range from 0 (no confounding) to 1 (entirely confounded). In the extreme case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$. In this scenario, the SNP-phenotype associations will be unbiased, but the residual variance will be inflated.

@@

Given the nature of environmental confounding in our setting of interest, and in order to assess methodological performance in terms of estimation accuracy and bias, we simulate from the data-generating model
\begin{equation}
    \label{eqn:sim}
    \by = a \bX \bbeta + b \bZ \bgamma + c \beps,
\end{equation}
where $\by$ represents the pseudophenotype, $\bgamma$ is a fixed quantity (see section~\ref{sec:sim_env_conf}), and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1 and is partitioned into genetic signal, environmental confounding, and random noise components,
\begin{equation}
    \label{eqn:parition_y}
    \mathbb{V}(\by) = \eta + (1 - \eta) \xi + (1-\eta)(1-\xi).
\end{equation}
$\eta$ defines the proportion of the variance of $\by$ attributable to additive genetic factors, commonly referred to as the narrow-sense heritability, $h^2 = \sigmagg/(\sigmagg + \sigmaee)$. $\eta$ also describes the data-generating model's signal to noise ratio (SNR). $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise. So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to unobserved environmental effects.

>>>>

Furthermore, the proportion of the environmental effect in the column space of $\boldsymbol{P}_{\bX}$ (i.e., the portion that will cause confounding) is given by
\begin{equation}
    \label{eqn:ratio}
    ||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2,
\end{equation}
which we refer to as the Confounding Proportion (CP). The CP can range from 0 (no confounding) to 1 (entirely confounded). In the extreme case where $\bZ \bgamma$ is completely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$. In this scenario, the SNP-phenotype associations will be unbiased, but the residual error will be inflated.

@@

Given the nature of environmental confounding in our setting of interest, and in order to assess methodological performance in terms of estimation accuracy and bias, we simulate from the data-generating model
\begin{equation}
    \label{eqn:sim}
    \by = a \bX \bbeta + b \bZ \bgamma + c \beps,
\end{equation}
where $\by$ represents the pseudophenotype, $\bgamma$ is a fixed quantity (see section~\ref{sec:sim_env_conf}), and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1 and is partitioned into genetic signal, environmental confounding, and random noise components,
\begin{equation}
    \label{eqn:parition_y}
    \mathbb{V}(\by) = \eta + (1 - \eta) \xi + (1-\eta)(1-\xi).
\end{equation}
$\eta$ defines the proportion of the variance of $\by$ attributable to additive genetic factors, commonly referred to as the narrow-sense heritability, $h^2 = \sigmagg/(\sigmagg + \sigmaee)$. The parameter $\eta$ also describes the data-generating model's signal to noise ratio (SNR), while $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise. So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to unobserved environmental effects.
